---
title: "SEM Strategies for COIL-related Data"
author: "Juan C. Correa"
date: "2024-02-01"
output: html_document
---

# Modeling Strategies

In this Supplementary Material, we will show two strategies of structural equation modeling. In our first strategy, we follow a "top-down modeling" which deviates from the second strategy called "bottom-up modeling" in several ways.

```{r}
knitr::kable(head(mtcars[, 1:4]), "pipe")
```




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
coildata <- coildata <- read_csv("Data/coildata.csv")
```

# Testing Univariate and Multivariate Normal Distribution Assumptions

The evaluation of these assumptions is relatively straightforward through exploratory data visualization of kernel density plots and the Henze-Zirkler test implemented in the R package MVN

```{r}
library(ggplot2)

coildata_long <- reshape2::melt(coildata)

ggplot(coildata_long, aes(x = value, fill = variable)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~variable, scales = "free", ncol = 5) +
  theme_minimal() +  # Optional: Customize the theme
  labs(title = "Statistical distributions of observed variables")

library(MVN)
mvn(coildata)
```

These results show that the univariate and multivariate normal distribution assumptions are unmet. As a result, these violations can be a possible lead to biased estimates and a poor fit when testing the psychometric structure of the transactive memory systems scale.

# Testing Unidimensional Structures

Now, we follow a bottom-up model specification according to which we are going to test the unidimensionality of the observed variables for the specialization sub-scale, the credibility sub-scale, and the coordination sub-scale. 

## Unidimensionality of Specialization
