---
title: "SEM Strategies for COIL-related Data"
author: "Juan C. Correa"
date: "2024-02-01"
output: html_document
---


In this Supplementary Material, we will show two strategies of structural equation modeling. In our first strategy, we follow a "top-down modeling" which deviates from the second strategy called "bottom-up modeling" in several ways.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
coildata <- coildata <- read_csv("Data/coildata.csv")
```

# Testing Univariate and Multivariate Normal Distribution Assumptions

The evaluation of these assumptions is relatively straightforward through exploratory data visualization of kernel density plots and the Henze-Zirkler test implemented in the R package MVN

```{r}
library(ggplot2)

coildata_long <- reshape2::melt(coildata)

ggplot(coildata_long, aes(x = value, fill = variable)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~variable, scales = "free", ncol = 5) +
  theme_minimal() +  # Optional: Customize the theme
  labs(title = "Statistical distributions of observed variables")

library(MVN)
mvn(coildata)
```

These results show that the univariate and multivariate normal distribution assumptions are unmet. As a result, these violations can be a possible lead to biased estimates and a poor fit when testing the psychometric structure of the transactive memory systems scale.

# Testing Unidimensional Structures

Now, we follow a bottom-up model specification according to which we are going to test the unidimensionality of the observed variables for the specialization sub-scale, the credibility sub-scale, and the coordination sub-scale. 

## Unidimensionality of Specialization

```{r, include=FALSE, echo=TRUE}
library(lavaan);
modelData <- coildata
 model<-"
! regressions 
   SP=~SP__SP1*SP1
   SP=~SP__SP2*SP2
   SP=~SP__SP3*SP3
   SP=~SP__SP4*SP4
   SP=~SP__SP5*SP5
! residuals, variances and covariances
   SP1 ~~ VAR_SP1*SP1
   SP2 ~~ VAR_SP2*SP2
   SP3 ~~ VAR_SP3*SP3
   SP4 ~~ VAR_SP4*SP4
   SP5 ~~ VAR_SP5*SP5
   SP ~~ 1.0*SP
! observed means
   SP1~1;
   SP2~1;
   SP3~1;
   SP4~1;
   SP5~1;
";
result <- cfa(model, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE)
summary(result, fit.measures=TRUE)
```

The results of this simple model are far from conclusive, because some of its Goodness of fit are ok (e.g., CFI = 1.00, RMSEA = 0.00) but others are not (e.g., SRMR = 0.12 and confidence intervals for RMSEA include values that exceed acceptable values 0.00 >= RMSEA <= 0.13). 

```{r}
SP <- summary(result, fit.measures=TRUE)
SP <- data.frame(SP$fit)
SP
```

The value of 0.12 for standardized root mean squared residual (SRMR) can be an indicator of some problems of misspecification error. We will check if this is the case by looking at the values of factor loadings between specialization and its five observed variables. The transactive memory systems scale proposed for English speakers by Lewis (2003) and validated for Spanish speakers by García-Chitiva (2021) indicate that all factor loadings should be positive and statistically significant. If this implied hypothesis is not met, specification error might be due to those factor loadings that behave differently.

```{r}
SP <- summary(result, fit.measures=TRUE)
knitr::kable(head(SP$pe, 5), "pipe")
```

As expected, all factor loadings are positive and highly significant except that of the second item (SP2) whose factor loading is almost zero and non-significant ($\lambda$ = 0.079, se = 0.17, p = 0.64). This particular result will lead us to reduce the complexity of this unidimensional model for Specialization by removing SP2 from the proposed structure (i.e., just four items). If this simpler model behaves as expected we can keep it for further analyses.

```{r}
 modelSP <-"
! regressions 
   SP=~SP__SP1*SP1
   SP=~SP__SP3*SP3
   SP=~SP__SP4*SP4
   SP=~SP__SP5*SP5
! residuals, variances and covariances
   SP1 ~~ VAR_SP1*SP1
   SP3 ~~ VAR_SP3*SP3
   SP4 ~~ VAR_SP4*SP4
   SP5 ~~ VAR_SP5*SP5
   SP ~~ 1.0*SP
! observed means
   SP1~1;
   SP3~1;
   SP4~1;
   SP5~1;
"
result2 <- cfa(modelSP, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE)

SP2 <- summary(result2, fit.measures=TRUE)
SP2 <- data.frame(SP2$fit)
SP2
```

As expected, this simpler model shows a better performance (CFI = 0.99; rmsea = 0.05; SRMR =  0.02) and all factor loadings are positive and highly significant.

```{r}
SP2 <- summary(result2, fit.measures=TRUE)
knitr::kable(head(SP2$pe, 4), "pipe")
```


## Unidimensionality of Credibility

We continue our analyses with the sub-scale of credibility. The results of this simple model are unsatisfactory due to its poor fit (e.g., CFI = 0.87, RMSEA = 0.22, SRMR = 0.27).


```{r, include=FALSE, echo=TRUE}
library(lavaan);
modelData <- coildata
model<-"
! regressions 
   CR=~CR__CR1*CR1
   CR=~CR__CR2*CR2
   CR=~CR__CR3*CR3
   CR=~CR__CR4*CR4
   CR=~CR__CR5*CR5
! residuals, variances and covariances
   CR1 ~~ VAR_CR1*CR1
   CR2 ~~ VAR_CR2*CR2
   CR3 ~~ VAR_CR3*CR3
   CR4 ~~ VAR_CR4*CR4
   CR5 ~~ VAR_CR5*CR5
   CR ~~ 1.0*CR
! observed means
   CR1~1;
   CR2~1;
   CR3~1;
   CR4~1;
   CR5~1;
";

result3 <- cfa(model, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE);
summary(result3, fit.measures=TRUE);
```

```{r}
CR <- summary(result3, fit.measures=TRUE)
CR <- data.frame(CR$fit)
CR
```

When looking at the factor loadings of this sub-scale, we find two items (i.e., CR4 and CR5) whose values are far from what we expect from theory because they showed non-significant negative values. 

```{r}
CR <- summary(result3, fit.measures=TRUE)
knitr::kable(head(CR$pe, 5), "pipe")
```

We then move on to reduce the complexity of this credibility sub-scale by specifying a structure with just three items (i.e., CR1, CR2, and CR3).

```{r}
modelData <- coildata
 model<-"
! regressions 
   CR=~CR__CR1*CR1
   CR=~CR__CR2*CR2
   CR=~CR__CR3*CR3
! residuals, variances and covariances
   CR1 ~~ VAR_CR1*CR1
   CR2 ~~ VAR_CR2*CR2
   CR3 ~~ VAR_CR3*CR3
   CR ~~ 1.0*CR
! observed means
   CR1~1;
   CR2~1;
   CR3~1;
";

result4 <- cfa(model, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE);

CR2 <- summary(result4, fit.measures=TRUE)
CR2 <- data.frame(CR2$fit)
CR2
```

```{r}
CR2 <- summary(result4, fit.measures=TRUE)
knitr::kable(head(CR2$pe, 3), "pipe")
```

## Unidimensionality of Coordination

Like previous sub-scale, let us show the results for the unidimensionality of coordination with all of its five observed variables. Here, the results are even worse than previous sub-scales, as the goodness-of-fit indexes reveal unsatisfactory values (i.e., CFI = 0.78; RMSEA = 0.27; SRMR = 0.12).  

```{r, include=FALSE, echo=TRUE}
library(lavaan);
modelData <- coildata
 model<-"
! regressions 
   CD=~CD__CD1*CD1
   CD=~CD__CD2*CD2
   CD=~CD__CD3*CD3
   CD=~CD__CD4*CD4
   CD=~CD__CD5*CD5
! residuals, variances and covariances
   CD1 ~~ VAR_CD1*CD1
   CD2 ~~ VAR_CD2*CD2
   CD3 ~~ VAR_CD3*CD3
   CD4 ~~ VAR_CD4*CD4
   CD5 ~~ VAR_CD5*CD5
   CD ~~ 1.0*CD
! observed means
   CD1~1;
   CD2~1;
   CD3~1;
   CD4~1;
   CD5~1;
";
result5 <- cfa(model, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE)
summary(result5, fit.measures=TRUE)
```

```{r}
CD <- summary(result5, fit.measures=TRUE)
CD <- data.frame(CD$fit)
CD
```

When looking at the factor loadings, it turns out that this sub-scale has two items (i.e., CD3 and CD5) whose values deviate from theoretical considerations as they show non-significant negative factor loadings. 

```{r}
CD <- summary(result5, fit.measures=TRUE)
knitr::kable(head(CD$pe, 5), "pipe")
```

We remove these two items and contrast a simpler model with the three items that showed the expected statistical behavior implied by the theory of transactive memory systems scale. Now, the goodness-of-fit indexes proved to be better (i.e., CFI = 1.00; RMSEA = 0.00; SRMR = 3.68e-08).

```{r}
library(lavaan);
modelData <- coildata
 model<-"
! regressions 
   CD=~CD__CD1*CD1
   CD=~CD__CD2*CD2
   CD=~CD__CD4*CD4
! residuals, variances and covariances
   CD1 ~~ VAR_CD1*CD1
   CD2 ~~ VAR_CD2*CD2
   CD4 ~~ VAR_CD4*CD4
   CD ~~ 1.0*CD
! observed means
   CD1~1;
   CD2~1;
   CD4~1;
";



result6 <- cfa(model, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE);

CD2 <- summary(result6, fit.measures=TRUE)
CD2 <- data.frame(CD2$fit)
CD2
```


```{r}
CD2 <- summary(result6, fit.measures=TRUE)
knitr::kable(head(CD2$pe, 3), "pipe")
```

# Testing Multidimensionality of Collaboration

When testing the unidimensionality of each sub-scale, we found evidence that none of them behave as expected with their five observed variables, and we had to test simpler models by removing those items with a gross deviation from theoretical considerartions. It goes without saying that if we test the multidimensionality of the transactive memory scale, its behavior will be far from satisfactory when all fifteen items are included in the specification. We will show that this is the case as follows.

```{r}
modelData <- coildata ;
modelCO <-"
! regressions 
   SP=~SP__SP1*SP1
   SP=~SP__SP2*SP2
   SP=~SP__SP3*SP3
   SP=~SP__SP4*SP4
   SP=~SP__SP5*SP5
   CR=~CR__CR1*CR1
   CR=~CR__CR2*CR2
   CR=~CR__CR3*CR3
   CR=~CR__CR4*CR4
   CR=~CR__CR5*CR5
   CD=~CD__CD1*CD1
   CD=~CD__CD2*CD2
   CD=~CD__CD3*CD3
   CD=~CD__CD4*CD4
   CD=~CD__CD5*CD5
! residuals, variances and covariances
   SP1 ~~ VAR_SP1*SP1
   SP2 ~~ VAR_SP2*SP2
   SP3 ~~ VAR_SP3*SP3
   SP4 ~~ VAR_SP4*SP4
   SP5 ~~ VAR_SP5*SP5
   CR1 ~~ VAR_CR1*CR1
   CR2 ~~ VAR_CR2*CR2
   CR3 ~~ VAR_CR3*CR3
   CR4 ~~ VAR_CR4*CR4
   CR5 ~~ VAR_CR5*CR5
   CD1 ~~ VAR_CD1*CD1
   CD2 ~~ VAR_CD2*CD2
   CD3 ~~ VAR_CD3*CD3
   CD4 ~~ VAR_CD4*CD4
   CD5 ~~ VAR_CD5*CD5
   SP ~~ 1.0*SP
   CR ~~ 1.0*CR
   CD ~~ 1.0*CD
   SP ~~ COV_SP_CR*CR
   CR ~~ COV_CR_CD*CD
   SP ~~ COV_SP_CD*CD
! observed means
   SP1~1;
   SP2~1;
   SP3~1;
   SP4~1;
   SP5~1;
   CR1~1;
   CR2~1;
   CR3~1;
   CR4~1;
   CR5~1;
   CD1~1;
   CD2~1;
   CD3~1;
   CD4~1;
   CD5~1;
"
result7 <- cfa(modelCO, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE);

CO <- summary(result7, fit.measures=TRUE)
knitr::kable(head(CO$pe, 15), "pipe")
CO$fit
```

The resulting goodness-of-fit indexes for this model revealed unacceptable values (i.e., CFI = 0.856, RMSEA = 0.127; SRMR = 0.109). Thus, we propose a simplified model that retains items that behaved as expected and discard those whose estimates grossly deviated from theoretical expectations. 

```{r, warning=FALSE}
modelData <- coildata ;
modelCO2 <-
"
! regressions 
   SP=~SP__SP1*SP1
   SP=~SP__SP3*SP3
   SP=~SP__SP4*SP4
   SP=~SP__SP5*SP5
   CR=~CR__CR1*CR1
   CR=~CR__CR2*CR2
   CR=~CR__CR3*CR3
   CD=~CD__CD1*CD1
   CD=~CD__CD2*CD2
   CD=~CD__CD4*CD4
! residuals, variances and covariances
   SP1 ~~ VAR_SP1*SP1
   SP3 ~~ VAR_SP3*SP3
   SP4 ~~ VAR_SP4*SP4
   SP5 ~~ VAR_SP5*SP5
   CR1 ~~ VAR_CR1*CR1
   CR2 ~~ VAR_CR2*CR2
   CD1 ~~ VAR_CD1*CD1
   CD2 ~~ VAR_CD2*CD2
   CD4 ~~ VAR_CD4*CD4
   SP ~~ 1.0*SP
   CR ~~ 1.0*CR
   CD ~~ 1.0*CD
   CR3 ~~ VAR_CR3*CR3
   SP ~~ COV_SP_CR*CR
   CR ~~ COV_CR_CD*CD
   SP ~~ COV_SP_CD*CD
! observed means
   SP1~1;
   SP3~1;
   SP4~1;
   SP5~1;
   CR1~1;
   CR2~1;
   CD1~1;
   CD2~1;
   CD4~1;
   CR3~1;
";

result8 <- cfa(modelCO2, data=coildata, fixed.x=FALSE, estimator="MLM", std.ov = TRUE, std.lv = TRUE);

CO2 <- summary(result8, fit.measures=TRUE)
ParEst2 <- data.frame(CO2$pe)
FIT2 <- data.frame(CO2$fit)

library(semPlot)
semPaths(result8, whatLabels = "std", layout = "tree", color = list(
  lat = c(rgb(0, 204, 0, maxColorValue = 255),
          rgb(255, 255, 0, maxColorValue = 255),
          rgb(255, 0, 0, maxColorValue = 255)),
  man = c(rgb(155, 253, 175, maxColorValue = 255),
          rgb(155, 255, 175, maxColorValue = 255),
          rgb(155, 255, 175, maxColorValue = 255),
          rgb(155, 255, 175, maxColorValue = 255),
          rgb(255, 255, 153, maxColorValue = 255),
          rgb(255, 255, 153, maxColorValue = 255),
          rgb(255, 255, 153, maxColorValue = 255),
          rgb(255, 102, 102, maxColorValue = 255),
          rgb(255, 102, 102, maxColorValue = 255),
          rgb(255, 102, 102, maxColorValue = 255))),
  edge.color = "black",
  edge.label.cex = 1,
  edge.width = 1.5,
  label.cex = 1,
  node.width = 1,
  node.height = 1,
  mar = c(5, 1, 5, 1), intercepts = FALSE, residuls = FALSE, nCharNodes = 0)
```

